\begin{thebibliography}{99}



\bibitem{kallwallace} Kall and Wallace [1994], \emph{Stochastic Programming, First Edition}, John Wiley and Sons.

\bibitem{bental} A. Ben-Tal and A. Nemirovski [1999], \emph{Robust
solutions of uncertain linear programs}, Operations Research Letters
25, (1999), 1-13.

\bibitem{rockafellar} R.T. Rockafellar and Roger J-B Wets [1991],
\emph{Scenarios and Policy Aggregation in Optimization Under
Uncertainty}, Mathematics of Operations Research, 16, (1991), 119-147.

\bibitem{kouvelis} Gutierrez and Kouvelis [1995], \emph{A robustness
approach to international sourcing}, Annals of Operations Research 59,
(1995), pp 165-193.

\bibitem{lin} Lin, Janak and Floudas [2004], \emph{A new robust
optimization approach for scheduling under uncertainty: I. Bounded
uncertainty}, Computers and Chemical Engineering 28, (2004),
pp 1069-1085.

\bibitem{tsutsui} Tsutsui and Ghosh [1997], \emph{Genetic Algorithms with a
robust solution searching scheme}, IEEE Transactions on Evolutionary
Computation, Vol. 1, No. 3. (September 1997).

\bibitem{spooren} Simon Spoorendonk [2008], \emph{Cut and Column Generation}, Ph.D. thesis, DIKU, (2008).

\bibitem{sandsynlighed} M. Sørensen [2006], \emph{En Introduktion til
  Sandsynlighedsregning, 7. udgave}, Afdeling for anvendt matematik og statistik.

\bibitem{statistik} I. Henningsen [2006], \emph{En Introduktion til
  Statistik Bind 1, 7. udgave}, Afdeling for anvendt matematik og statistik.

\bibitem{cormen} Cormen, Leiserson, Rivest and Stein [2001],
\emph{Introduction to Algorithms, 2nd edition}, MIT press, (2001).

\bibitem{russel} Russel and Norvig [2002], \emph{Artificial
Intelligence A Modern Approach, 2nd edition}, Prentice Hall, (2002).

\bibitem{wolsey} Laurence A. Wolsey [1998], \emph{Integer programming,
1st edition}, Wiley-Interscience, (1998).

\bibitem{taha} H. A. Taha [2002], \emph{Operations Research - An
  Introduction, 7th edition}, Prentice Hall, (2001).

\bibitem{bertsimas} D. Bertsimas and M. Sim [2004], \emph{The Price of
Robustness}, Operations Research, Vol. 52, No. 1, (January-February
2004), pp. 35-53.

\bibitem{tablesandmiller} http://www.core.ucl.ac.be/wolsey/lotsizel.htm

\end{thebibliography}